Mixing time estimation in reversible Markov chains from a single sample path

### 2015-06-02 rates.tex
minor

\todoc{Turn this paragraph into a remark?}
Note that because the eigenvalues of $\vL$ are the same as that of the
transition probability matrix $\vP$, 
we could have instead opted to
estimate $\vP$, say, using simple frequency estimates obtained from
the sample path, and then computing the second largest eigenvalue of
this empirical estimate $\wh\vP$.
The difficulty with this approach is that $\vP$ lacks the structure
required by certain strong eigenvalue perturbation results.
One could instead invoke the Ostrowski-Elsner theorem
\citep[cf.~Theorem 1.4 on Page 170 of][]{stewart1990matrix}, which
bounds the \emph{matching distance} between the eigenvalues of a
matrix $\vA$ and its perturbation $\vA+\vE$ by $O(\norm{\vE}^{1/d})$.
%\footnote{Actually, this result is not even readily applicable: 
%the Haussdorff distance is actually an optimistic estimate of the deviation between
%$\gap$ and its estimate.}
Since $\norm{\wh\vP-\vP}$ is expected to be of size $O(n^{-1/2})$,
this approach will give a confidence interval for $\gap$ whose width
shrinks at a rate of $O(n^{-1/(2d)})$---an exponential slow-down
compared to the rate from \cref{thm:err}.
As demonstrated through an example from \citet{stewart1990matrix}, the
dependence on the $d$-th root of the norm of the perturbation cannot
be avoided in general.
Our approach based on estimating a symmetric matrix affords us the use
of perturbation results that exploit more structure.
%(In \cref{sec:empirical}, we develop another estimator of
%$\vL$---in order to exploit symmetry---that is based on estimating
%$\vP$ for reasons discussed below.)

@@

\subsection{Lower bounds}

>>>>

\todoc{Turn this paragraph into a remark?}
Note that because the eigenvalues of $\vL$ are the same as that of the
transition probability matrix $\vP$, 
we could have instead opted to
estimate $\vP$, say, using simple frequency estimates obtained from
the sample path, and then computing the second largest eigenvalue of
this empirical estimate $\wh\vP$.
(In fact, this would allow the results to extend to non-reversible
chains.)
The difficulty with this approach is that $\vP$ lacks the structure
required by certain strong eigenvalue perturbation results.
One could instead invoke the Ostrowski-Elsner theorem
\citep[cf.~Theorem 1.4 on Page 170 of][]{stewart1990matrix}, which
bounds the \emph{matching distance} between the eigenvalues of a
matrix $\vA$ and its perturbation $\vA+\vE$ by $O(\norm{\vE}^{1/d})$.
%\footnote{Actually, this result is not even readily applicable: 
%the Haussdorff distance is actually an optimistic estimate of the deviation between
%$\gap$ and its estimate.}
Since $\norm{\wh\vP-\vP}$ is expected to be of size $O(n^{-1/2})$,
this approach will give a confidence interval for $\gap$ whose width
shrinks at a rate of $O(n^{-1/(2d)})$---an exponential slow-down
compared to the rate from \cref{thm:err}.
As demonstrated through an example from \citet{stewart1990matrix}, the
dependence on the $d$-th root of the norm of the perturbation cannot
be avoided in general.
Our approach based on estimating a symmetric matrix affords us the use
of perturbation results that exploit more structure.
%(In \cref{sec:empirical}, we develop another estimator of
%$\vL$---in order to exploit symmetry---that is based on estimating
%$\vP$ for reasons discussed below.)

@@

\subsection{Lower bounds}
\label{sec:rates-lower}
